---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

# Load packages
```{r}
library(jsonlite)
library(dplyr)
library(crosstable)
library(ggplot2)
library(nlme)
library(texreg)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(coefplot)
library(jtools)
library(ggstance)
library(tidyr)
library(usmap) #import the package
library(ggplot2) #use ggplot2 to add layer for visualization
library(mapview)
library(sf)
library(PNWColors)
library(transformr)
library(readr)
library(stringr)
options(dplyr.summarise.inform = FALSE)
```

# Upload USPTO data and data treatment
```{r}
# upload the patent data
pharma_patent <- read.csv("/project/biocomplexity/sdad/projects_data/uspto/PatentsView/working/pharma_pat_inv_loc.csv")
pharma_patent$field <- "pharmaceuticals"
civil_patent<- read.csv("/project/biocomplexity/sdad/projects_data/uspto/PatentsView/working/civil_pat_inv_loc.csv")
civil_patent$field <- "civil engineering"

# combine data from the two fields
patent <- dplyr::bind_rows(pharma_patent, civil_patent)
patent$female_flag <- 1-patent$male_flag
patent$gender[patent$male_flag==1] <- "male"
patent$gender[patent$male_flag==0] <- "female"
patent <- patent %>% mutate(gender = ifelse(is.na(gender), 'unknown', gender))

# subset to US, count the number of patent by inventor over time
US_patent <- patent %>% filter(country=="US") 
```

# Descriptive analysis

## Gender distribution among US inventors
```{r}
# gender distribution by fields
US_patent_desc <- US_patent %>% 
  select(inventor_id,gender,field) %>% 
  unique()

crosstable(US_patent_desc, gender, by=field, total= "all", margin="column")
```

There is more women inventor in Pharma (22.54%) than Civil engineering (6.76%).


## Trends in the gender proportion of inventors, number of patents by gender
```{r}
US_patent01 <- US_patent %>% 
  select(inventor_id,patent_id,gender,year,field) %>%
  group_by(gender,year,field) %>% 
  mutate(Ninventors = length(unique(inventor_id)), 
         Npatents = length(unique(patent_id)),
         .keep = "unused") %>%
  unique()

# compute the female proportion by field over time
US_patent02 <- US_patent01 %>% group_by(year,field) %>% summarise(total_inventors=sum(Ninventors), total_patents=sum(Npatents))
US_patent02 <- merge(US_patent01, US_patent02, by=c('year','field'))
US_patent02$prop_gender <- 100*US_patent02$Ninventors/US_patent02$total_inventors

# plot the trends
ggplot(US_patent02[US_patent02$gender=='female',], aes(x=year, y=Ninventors)) + geom_line(aes(colour=field)) + labs(x = "year", y="Number of women inventors", title='number of women inventors in US')
ggplot(US_patent02[US_patent02$gender=='female',], aes(x=year, y=prop_gender)) + geom_line(aes(colour=field)) + labs(x = "year", y="proportion of women", title='proportion of women inventors in US')

```


## growth rate difference
```{r}
reg01 <- lm(log(prop_gender) ~ year + factor(field)*year -1, data = US_patent02[US_patent02$gender=='female',])
summary(reg01)
```

## collaboration between women inventors
```{r}
# proportion of female by patent
US_patent_gender <- US_patent %>% group_by(patent_id,field,year) %>% summarise(patent_diversity_rate=100*mean(1-male_flag,na.rm=T))

# for patents with at least on female, what is the average female propotion over time and field
US_patent_gender$dum_diverse <- US_patent_gender$patent_diversity_rate
US_patent_gender$dum_diverse[US_patent_gender$patent_diversity_rate==0] <- NA
patent_diversity <- US_patent_gender %>% 
  group_by(field,year) %>% 
  summarise(diversity_rate=100*mean(patent_diversity_rate>0, na.rm=T), score_diversity=mean(dum_diverse, na.rm=T))

# plot
par(mfrow=c(1,2)) 
ggplot(patent_diversity, aes(x=year, y=diversity_rate)) + geom_line(aes(colour=field)) + labs(x = "year", y="proportion of patents", title='proportion of patents with at least one women inventors over time')
ggplot(patent_diversity, aes(x=year, y=score_diversity)) + geom_line(aes(colour=field)) + labs(x = "year", y="average proportion of women", title='average proportion of women inventors by patents over time')

```

The trends in the porportion of patent with at least one female is similar than the proportion of women inventors. In addition, the average proportion of women by patent is decreasing (even more in pharma). This suggest that inventors collaboration is not driven th 

```{r}
hist(US_patent_gender$patent_diversity_rate)
```


